---
title: "The Philip Curve & Forecasting Inflation"
output: html_notebook
---
Load the libraries
```{r}
library(tidyverse)
library(mosaic)
```
Data
```{r}
data("USMacroSW", package = "AER")
```
#The Philip Curve
Calculate the quarterly inflation rate by taking the difference of log CPI.
```{r}
usm <- ts.intersect(USMacroSW, 4 * 100 * diff(log(USMacroSW[,"cpi"])))
```
Adding column names
```{r}
colnames(usm) <- c((colnames(USMacroSW)), "infl")
```
Extracting umemployment & inflation rate data and divide them into before and after 1970
```{r}
unemp1 <- usm[, "unemp"]
unemp_b <- unemp1[time(unemp1) < 1970]
unemp_a <- unemp1[time(unemp1) >= 1970]
infl1 <- usm[, "infl"]
infl_b <- infl1[time(infl1) < 1970]
infl_a <- infl1[time(infl1) >= 1970]
```
Plot inflation rate vs. unemployment before and after 1970
```{r}
xyplot(infl_b ~ unemp_b, type = c("p", "smooth"))
xyplot(infl_a ~ unemp_a, type = c("p", "smooth"))
```
#Forecasting Inflation
Plot the inflation rate and its Autocorrelation Function
```{r}
plot(infl1, las = 1)
acf(infl1, main = "")
```
Load the "tseries" package
```{r}
library(tseries)
```
Use the Augmented Dickey-Fuller Test to test the null hypothesis that an unit root is present in the time series. In other words, to test if β1 = 1 in the equation Yt = β0 + β1Yt−1 + ut .
```{r}
adf.test(usm[,"infl"])
```
The p-value indicates that the null hypothesis is not rejected. Plot and test again using the difference of inflation rate.
```{r}
diff.infl <- diff(usm[, "infl"])
plot(diff.infl, las = 1)
acf(diff.infl, main = "")
adf.test(diff(usm[,"infl"]))
```
The null hypothesis is rejected.

Autoregressive model with past values:
Zt = β0 + β1Zt−1 + ut
Predictive model:
Zt+1 = β0<EST> + β1<EST> + Zt ,

Load the "forecast" package
```{r}
library(forecast)
```
Calculating, printing and plotting the forecast data
```{r}
fit_ar4 <- Arima(infl1, order = c(2, 1, 0))
forecast(fit_ar4, level = 95)
plot(forecast(fit_ar4), las = 1, main = "")
```

